---
title: "Advanced R"
subtitle: "Writing Faster R Code"
author: "Martin Schnürch & David Izydorczyk"
date: "17.06.2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "default-fonts","subfiles/my-theme.css","subfiles/my-fonts.css","xaringan-themer.css"]
    nature:
      highlightStyle: github
      ratio: "16:9"
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, message = FALSE, warning=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(tictoc)
library(microbenchmark)
library(benchmarkme)
library(bench)
library(data.table)
library(kableExtra)
library(foreach)
```


class: heading,middle


Part 3: Parallelization with `foreach`



<!-- *********** NEW SLIDE ************** -->
---
## foreach

- `foreach` is like a for loop, but is implemented using a binary operator, called `%do%`.
- The main difference between `foreach` and an old boring for loop is that it returns a value (you can also specify the output format)
- ! `i = 1:5` instead of `i in 1:5`

.pull-left[

```{r}
for(i in  1:2){
  sqrt(i)
}
```

```{r}
library(foreach)

foreach(i = 1:2,.combine = "list") %do% {
  sqrt(i)
}
```

] 

.pull-right[

```{r}
library(foreach)

foreach(i = 1:2,.combine = "c") %do% {
  sqrt(i)
}
```

```{r}
library(foreach)

foreach(i = 1:2,.combine = "+") %do% {
  sqrt(i)
}
```

]


<!-- *********** NEW SLIDE ************** -->
---
## foreach

- The real power of `foreach` is when you combine it with `doSNOW` (or `doParallel`)
- The `doSNOW` package is a “parallel backend” for the `foreach` package


```{r eval=FALSE}

cl <- makeSOCKcluster(ncores)  # define cluster
registerDoSNOW(cl) # start cluster 

res <- foreach(iterator = 1:100,  #define iterator
              .combine  = "cbind", # define how the output should be combined
              .packages = c("tidyverse"), #  packages which should be loaded on the workers 
              ...
              ) %dopar% {
                   
                   # your function goes here 
                   
       }
  
stopCluster(cl) # stop cluster

```




<!-- *********** NEW SLIDE ************** -->
---
## Example

.pull-left[


```{r eval=FALSE}
boot_f <- function(nreps=1e4){
  
  n  <- nrow(mtcars)
  rs <- vector(mode="integer",length=nreps)
  
  for(reps in 1:nreps){
    
    ind  <- sample(n, n, replace=TRUE)
    temp <- mtcars[ind,]
    mod  <- lm(mpg~wt+disp,data=temp)
    
    
    rs[reps] <- summary(mod)$r.square

  }
  
  return(rs)
}

```


]

.pull-right[

```{r eval=FALSE}
run_sim  <- function(ncores,nreps) {
  
cl <- makeSOCKcluster(ncores) #<<
registerDoSNOW(cl)#<<

res <- foreach(i = 1:100,
              .combine = "cbind") %dopar% {
                   
                   boot_f(nreps)
                   
       }
  
stopCluster(cl)#<<
}
```

]

<!-- *********** NEW SLIDE ************** -->
---
## Example

```{r eval=FALSE}
library(doSNOW)
library(foreach)
library(bench)

results <- bench::press(
            ncores = c(1,4,8,32,60),
            nreps  = c(10,1e3,1e4,1e5),
            {
              bench::mark(
                sim = run_sim(ncores,nreps)
              )
            }
          )
```



Cores             |  10 reps  |  1,000 reps  |   10,000 reps  |  100,000 reps | 
------------------|-----------|--------------|----------------|---------------|
1 Core            |1.08 s     |   1.14   m   | 11.34 m        |     1.86 h    |
4 Cores           |1.02 s     |   19.30  s   | 3.06 m         |    30.11 m    |
8 Cores           |17.1 s     |    12.12 s   | 1.67 m         |    16.65 m    |
32 Cores          |5.89 s     |   10.42  s   | 51.02 s        |     7.29 m    |
60 Cores          |10.98 s    |   16.22  s   | 57.27 s        |     7.24 m    |



<!-- *********** NEW SLIDE ************** -->
---
## Foreach

Each iteration should execute computationally-intensive work. Scheduling tasks has overhead, and can exceed the time to complete the work itself for small jobs.

<!-- *********** NEW SLIDE ************** -->
---
## Adding a progress bar

You can also add a progress bar (very useful)

```{r eval=FALSE}
iterations <- 8

ncores     <- 4
cl         <- makeSOCKcluster(ncores)
  
pb         <- txtProgressBar(max = iterations, style = 3)#<<
progress   <- function(n){setTxtProgressBar(pb, n)}#<<
opts       <- list(progress = progress)#<<

registerDoSNOW(cl)


res <- foreach(i = 1:iterations,
              .combine       = list,
              .options.snow  = opts #<<
             ) %dopar% {
  
               
               
                  boot_f(nreps=5000)
  
              }

stopCluster(cl)
```




<!-- *********** NEW SLIDE ************** -->
---
class:small
## Nested foreach


- You can also use nested `foreach`-loops
- Example use case: Loop over a simulation design matrix and then repetitions 



```{r message=FALSE, warning=FALSE}
library(doSNOW)
library(foreach)

# Make Matrix
m <- matrix(1:2500,nrow=50,ncol=50)

# Start Cluster
cl       <- makeSOCKcluster(4)
registerDoSNOW(cl)


# Run Foreach
res <- foreach(col      = 1:ncol(m),
               .combine = "cbind"
               )  %:% #<<
          foreach(row      = 1:nrow(m),
                  .combine = "c") %dopar% {
  
                    m[col,row]^2
          
}

# Stop Cluster
stopCluster(cl)

# Show Results
#head(res)

```






<!-- *********** NEW SLIDE ************** -->
---
## Rcpp in foreach

- You can also use `Rcpp`-functions in `foreach`

```{r eval = FALSE}

## Do not run


cl       <- makeSOCKcluster(4)
registerDoSNOW(cl)

res <- foreach(iter     = 1:10,
               .noexport=c("your_super_cool_Rcpp"). #<<
               .combine = "rbind")   %dopar% {
  
          source("Scripts/functions/rcpp_functions.R") #<<
          
                 
          your_super_cool_Rcpp()
          
}


stopCluster(cl)



```




<!-- *********** NEW SLIDE ************** -->
---
## Your turn !

- Open the file `foreach_exercise.R`
- There you find a function to compute the power of a t-test via simulation using two for-loops
- Your task is it to make the same function with `foreach`






<!-- *********** NEW SLIDE ************** -->
---
## Helpful references:

- http://adv-r.had.co.nz/
- [foreach intro](https://cran.r-project.org/web/packages/doParallel/vignettes/gettingstartedParallel.pdf)
- [foreach intro 2](https://privefl.github.io/blog/a-guide-to-parallelism-in-r/)

